A cascaded supervised learning approach to inverse reinforcement learning

### 2013-04-05 piot2012cascading_ECML/Cascading.tex
Version ECML sur 16 pages.

%\usepackage{amsthm}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{soul}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage[integrals]{wasysym}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{subfigure}
%\usepackage{natbib}
% scientific notation, 1\e{9} will print as 1x10^9
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}
\usepackage{booktabs} % Fancy tables
\tolerance=1000
\usepackage{amsmath}
%\newtheorem{definition}{Definition}
%\newtheorem{theorem}{Theorem}
%\newtheorem{corollary}{Corollary}
\usepackage{dsfont}
\usepackage{mathtools}
\mathtoolsset{showonlyrefs=true}
\providecommand{\alert}[1]{\textbf{#1}}
\newcommand{\E}{\mathbf{E}}
\newcommand{\argmax}{\operatorname*{argmax}} %\operatorname* pour les op. pouvant admettre des limites...
\newcommand{\Card}{\operatorname*{Card}} %\operatorname* pour les op. pouvant admettre des limites...

@@

We work with finite MDPs \cite{puterman1994markov}, that is tuples $\{S,A,P,R,\gamma\}$. The state space is noted $S$, $A$ is a finite action space, $R\in\mathbb{R}^{S\times A}$ is a reward-function, $\gamma\in (0,1)$ is a discount factor and $P\in \Delta_{S}^{S\times A}$ is the Markovian dynamics of the MDP. Thus, for each $(s,a)\in S\times A$, $P(.|s,a)$ is a distribution over $S$ and $P(s'|s,a)$ is the probability to reach $s'$ by choosing action $a$ in state $s$. At each time step $t$, the agent uses the information encoded in the state $s_t\in S$ in order to choose an action $a_t \in A$ according to a (deterministic\footnote{We restrict ourselves here to deterministic policies, but the loss of generality is minimal as there exists at least one optimal deterministic policy.}) policy $\pi\in A^S$. The agent then steps to a new state $s_{t+1}\in S$ according to the Markovian transition probabilities $P(s_{t+1}|s_t,a_t)$. Given that $P_\pi = (P(s'|s,\pi(s)))_{s,s' \in S}$ is the irreductible transition probability matrix, the stationary distribution over the states $\rho_\pi$ induced by a policy $\pi$ satisfies
\begin{equation}
  \rho_\pi^TP_\pi = \rho_\pi^T\textrm{, with }X^T\textrm{ being the transpose of }X.
  \end{equation}
The stationary distribution relative to the exert policy $\pi_E$ is $\rho_E$.

The reward function $R$ is a local measure of the quality of the control. The global quality of the control induced by a policy $\pi$, with respect to a reward $R$, is assessed by the value function $V^\pi_R \in \mathbb{R}^{S}$ which associates to each state the expected discounted cumulative reward for following policy $\pi$ from this state:
\begin{equation}
V^\pi_R(s) = \E[\sum_{t\geq 0}\gamma^tR(s_t,\pi(s_t))|s_0 = s,\pi].
\end{equation}
This long-term criterion is what is being optimized when solving an MDP. Therefore, an optimal policy $\pi^*_R$ is a policy whose value function (the optimal value function $V^*_R$) is greater than that of any other policy, for all states: $\forall \pi, V^*_R\geq V^\pi_R$.

The Bellman evaluation operator $T^\pi_R: \mathbb{R}^{S} \rightarrow  \mathbb{R}^{S}$ is defined by
\begin{equation}
  T^{\pi}_RV = R_\pi + \gamma P_\pi V
  \end{equation}
where $R_\pi = (R(s,\pi(s)))_{s\in S}$. The Bellman optimality operator follows naturally:
\begin{equation}
  T^*_RV = \max_\pi T^\pi_RV.
\end{equation}
  Both operators are contractions. The fixed point of the Bellman evaluation operator $T^\pi_R$ is the value function of $\pi$ with respect to reward R:
  \begin{eqnarray}
    V^\pi_R &=& T^\pi_R V^\pi_R \Leftrightarrow V^\pi_R = R_\pi + \gamma P_\pi V^\pi_R.
  \end{eqnarray}
The Bellman optimality operator $T^*_R$ also admits a fixed point, the optimal value function $V_R^*$ with respect to reward $R$.

@@

In this section, we prove that the deterministic expert policy $\pi_E$ is near optimal for the reward $\hat{R}^C$ the regression step outputs.
More formally, recalling from Sec.~\ref{sec:background} that $\rho_E$ is the stationary distribution of the expert policy, we prove that $\E_{s\sim\rho_E}[V^*_{\hat{R}^C}(s)-V^{\pi_E}_{\hat{R}^C}(s)]$ is bounded by a term that depends on:
\begin{itemize}
  \item the classification error defined as
\begin{equation}  
\epsilon_C=\E_{s\sim\rho_E}[\mathds{1}_{\{\pi_C(s)\neq\pi_E(s)\}}];
\end{equation}
\item the regression error defined as 
  \begin{equation}
    \epsilon_R=\max_{\pi\in A^S}\|\epsilon_\pi^R\|_{1,\rho_E},
  \end{equation}
  with:
  \begin{itemize}
  \item the subscript notation already used for $R_\pi$ and $P_\pi$ in Sec.~\ref{sec:background} meaning that, given an $X\in\mathbb{R}^{S\times A}$, $\pi\in A^S$, and $a\in A$,  $X_\pi\in\mathbb{R}^S$ and $X_a\in\mathbb{R}^S$ are respectively such that: $\forall s\in S, X_\pi(s)=X(s,\pi(s))$ and $\forall s\in S, X_a(s)=X(s,a)$ ;
  \item $\epsilon^R_\pi=R^C_\pi-\hat{R}^C_\pi$ ;
  \item $\|.\|_{1,\mu}$ the $\mu$-weighted $L_1$ norm: $\|f\|_{1,\mu} = \E_{x\sim \mu}[|f(x)|]$;
  \end{itemize}
\item the concentration coefficient $C_* = C_{\hat \pi_C}$ with:
  \begin{itemize}
  \item $C_{\pi}=(1-\gamma)\sum_{t\geq0}\gamma^tc_{\pi}(t), \text{ with } c_{\pi}(t)=\max_{s\in S}\frac{(\rho_E^TP^t_\pi)(s)}{\rho_E(s)}$ ;
  \item $\hat \pi_C$, the optimal policy for the reward $\hat R^C$ output by the algorithm ;
  \end{itemize}
  The constant $C_*$ can be estimated \emph{a posteriori} (after $\hat{R}^C$ is computed). {\it A priori}, $C_*$ can be upper-bounded by a more usual and general concentration coefficient but $C_*$ gives a tighter final result: one can informally see $C_*$ as a measure of the similarity between the distributions induced by $\hat\pi_C$ and $\pi_E$ (roughly, if $\hat{\pi}_C\approx\pi_E$ then $C_* \approx 1$).
\item $\Delta q$, defined by 
  \begin{eqnarray}
    \Delta q &= &\max_{s\in S}(\max_{a\in A}q(s,a)-\min_{a\in A}q(s,a))=\max_{s\in S}(q(s,\pi_C(s))-\min_{a\in A}q(s,a)),
    \end{eqnarray} which could be normalized to $1$ without loss of generality. The range of variation of $q$ induces the one of $R^C$, $\hat{R}^C$ and $V^{\pi_C}_{\hat{R^C}}$.
  \end{itemize}
\begin{theorem}
\label{thm}
Let $\pi_E$ be the deterministic expert policy, $\rho_E$ its stationary distribution and $\hat{R}^C$ the reward the cascading algorithm outputs. We have:
\begin{equation}
0\leq\E_{s\sim\rho_E}[V^*_{\hat{R}^C}(s)-V^{\pi_E}_{\hat{R}^C}(s)]\leq \frac{1}{1-\gamma}\left(\epsilon_C\Delta q +\epsilon_R(1+C_*)\right).
\end{equation}
\end{theorem}
\begin{proof}
First let's recall some notations, $q\in\mathbb{R}^{S\times A}$ is the score function output by the classification step, $\pi_C$ is a deterministic classifier policy so that $\forall s \in S,\pi_C(s)\in\argmax_{a\in A}q(s,a)$, $R^C\in\mathbb{R}^{S\times A}$ is:
\begin{equation}
\forall (s,a)\in S\times A, R^C(s,a)=q(s,a) -\gamma\sum_{s'\in S}P(s'|s,a)q(s',\pi_C(s')),
\end{equation}
and $\hat{R}^C\in\mathbb{R}^{S\times A}$ is the reward function output by the regression step.

@@

We can now introduce the following vectorial equations:
\begin{align}
&R^C_a=q_a-\gamma P_aq_{\pi_C},
\\
&R^E_a=q_a-\gamma P_aq_{\pi_E},
\\
&\epsilon^R_a=R^C_a-\hat{R}^C_a.
\end{align}
Now, we are going to upper bound the term: $\E_{s\sim\rho_E}[V^*_{\hat{R}^C}-V^{\pi_E}_{\hat{R}^C}]\geq0$ (the lower bound is obvious as $V^*$ is optimal).
Recall that $\hat{\pi}_C$ is a deterministic optimal policy of the reward $\hat{R}^C$. First, the term $V^*_{\hat{R}^C}-V^{\pi_E}_{\hat{R}^C}$ is decomposed:
\begin{equation}
V^*_{\hat{R}^C}-V^{\pi_E}_{\hat{R}^C}=(V^{\hat{\pi}_C}_{\hat{R}^C}-V^{\hat{\pi}_C}_{R^C})+(V^{\hat{\pi}_C}_{R^C}-V^{\pi_E}_{R^C})+(V^{\pi_E}_{R^C}-V^{\pi_E}_{\hat{R}^C}).
\end{equation}
We are going to bound each of these three terms. First, let $\pi$ be a given deterministic policy. We have, using $\epsilon^R = R^C - \hat R^C$:
\begin{equation}
V^{\pi}_{R^C}-V^{\pi}_{\hat{R}^C}=V^{\pi}_{\epsilon^R}=(I-\gamma P_\pi)^{-1}\epsilon^R_{\pi}.
\end{equation}
If $\pi=\pi_E$, we have, thanks to the power series expression of $(I-\gamma P_{\pi_E})^{-1}$, the definition of $\rho_E$ and the definition of the $\mu$-weighted $L_1$ norm, one property of which is that $\forall X, \mu^TX \leq \|X\|_{1,\mu}$:
\begin{equation}
  \rho_E^T(V^{\pi}_{R^C}-V^{\pi}_{\hat{R}^C})=\rho_E^T(I-\gamma P_{\pi_E})^{-1}\epsilon^R_{\pi_E}=\frac{1}{1-\gamma}\rho_E^T\epsilon^R_{\pi_E}\leq\frac{1}{1-\gamma}\|\epsilon^R_{\pi_E}\|_{1,\rho_E}.\end{equation}

@@

\rho_E^T((V^{\pi_E}_{R^C}-V^{\pi_E}_{\hat{R}^C})+(V^{\hat{\pi}_C}_{\hat{R}^C}-V^{\hat{\pi}_C}_{R^C}))\leq\frac{1+C_*}{1-\gamma}\epsilon_R.

@@

\rho_E^T(V^{\pi_C}_{R^C}-V^{\pi_E}_{R^E})\leq\Delta q\sum_{s\in S}\rho_E(s)[\mathds{1}_{\{\pi_C(s)\neq\pi_E(s)\}}]= \Delta q \epsilon_C.

@@

\rho_E^T(V^{\pi_E}_{R^E}-V^{\pi_E}_{R^C})&=\rho_E^T(I-\gamma P_{\pi_E})^{-1}(R^E_{\pi_E}-R^C_{\pi_E}),
&=\rho_E^T(I-\gamma P_{\pi_E})^{-1}\gamma P_{\pi_E}(q_{\pi_C}-q_{\pi_E}),

@@

If we combine all of the results, we obtain the final bound:
0\leq\E_{s\sim\rho_E}[V^*_{\hat{R}^C}(s)-V^{\pi_E}_{\hat{R}^C}(s)]\leq \frac{1}{1-\gamma}(\epsilon_C\Delta q +\epsilon_R(1+C_*))).

@@

More precisely, the expert controls the system for $M$ runs of lengths $\{L_i\}_{1\leq i \leq M}$, %with values of $n$ in $\{3,7,10,15,20\}$,
giving samples $\{(s_k,a_k=\pi_E(a_k),s'_k)_k\} = D_E$. The dataset $D_C$ fed to the classifier is straightforwardly constructed from $D_E$ by dropping the $s'_k$ terms:
\begin{equation}
  D_C = \{(s_i = s_k,a_i=a_k)_i\}.
\end{equation}
\subsubsection{Heuristics}
\label{subsubsec:heuristics}

There is no particular problem for the classification step. It is however not reasonable to construct the dataset $D_R = \{((s_k,a_k=\pi_E(s_k)),\hat r_k)_k\}$ only from expert transitions and expect a small regression error term $\epsilon_R$. Indeed, the dataset $D_E$ only samples the dynamics induced by the expert's policy and not the whole dynamics of the MDP. 
This means that for a certain state $s_k$ we only know the corresponding expert action $a_k=\pi_E(s_k)$ and the following state $s'_k$ sampled according to the MDP dynamics : $s'_k \sim P(\cdot|s_k,a_k)$. For the regression to be meaningful, we need samples associating the same state $s_k$ and a different action $a\neq a_k$ with a datapoint $\hat r \neq \hat r_k$.\\
Recall that $\hat r_j = q(s_j,a_j) - \gamma q(s'_j,\pi_C(s'_j))$ (Eq.~\eqref{eq:rj}); without knowing $s'_k \sim P(\cdot|s_k,a\neq a_k)$, we cannot provide the regressor with a datapoint to associate with $(s_k,a\neq a_k)$.
We artificially augment the dataset $D_R$ with samples $((s_j=s_k,a),r_{min})_{j;\forall a\neq \pi_E(s_j) = a_k}$ where
\begin{equation}
  r_{min} = \min_k\hat r_k - 1.
\end{equation}
This heuristics intructs the regressor to associate a state-action tuple disagreeing with the expert ({\it i.e.}, $(s_k,a\neq a_k)$) with a reward strictly inferior to any of those associated with expert state action tuples ({\it i.e.}, $(s_k,a_k=\pi_E(s_k))$). Semantically, we are asserting that disagreeing with the expert in states the expert visited is a bad idea. This heuristics says nothing about states absent from the expert dataset. For such states the generalization capabilities of the regressor and, later on, the exploration of the MDP by an agent optimizing the reward will solve the problem.
Although this heuristics was not analyzed in Sec.~\ref{sec:analysis} (where the availability of a more complete dataset $D_R$ was assumed), the results shown in the next two subsections demonstrate its soundness.
\subsubsection{Comparison with state-of-the-art approaches}
The similar looking yet fundamentally different algorithm SCIRL \cite{klein2012scirl} is an obvious choice as a contender to CSI as it advertises the same ability to work with very little data, without repeatedly solving the RL problem. In both experiments we give the exact same data to CSI and SCIRL.

@@

The objective criterion for success is the number of steps needed to attain the goal when starting from a state picked at random ; the lesser the better. We can see Fig.~\ref{fig:MountainCar} that the optimal policies for the rewards found by SCIRL and CSI very rapidly attain expert-level performance and outperform the optimal policy for the reward of RE and the classification policy. When very few samples are available, CSI does better than SCIRL (with such a low $p$-value for $n=10$, see Tab.~\ref{tab:mc}, the hypothesis that the mean performance is equal can be rejected); SCIRL catches up when more samples are available. Furthermore, CSI required very little engineering as we cascaded two off-the-shelf implementations whereas SCIRL used hand-tuned features and a custom classifier.
\begin{table}
	\begin{center}
		\begin{tabular}{cc}
		\toprule
			Number of expert samples & $p$-value\\
			\midrule
                        $ 10$ & $\mathbf{1.5e-12}$\\
                        $ 30$ & $3.8e-01$\\
                        $100$ & $1.3e-02$\\
                        $300$ & $7.4e-01$\\
                        \bottomrule
		\end{tabular}
	\end{center}
        \caption{Student or Welch test of mean equality (depending on wether a Bartlett test of variance equality succeeds) $p$-values for CSI and SCIRL on the mountain car. High values ($>1.0\e{-02}$) means that the hypothesis that the means are equal cannot be rejected. As highlighted here, the means are different when the algorithms are fed only $10$ expert samples in which case CSI outperforms SCIRL.}
        \label{tab:mc}
\end{table}
\begin{figure}
  \includegraphics[width=\textwidth]{Exp11.pdf}
  \caption{Performance of various policies on the mountain car problem. This is the mean over 100 runs.}
  \label{fig:MountainCar}
\end{figure}

@@

Results are shown Fig.~\ref{fig:Highway}. We give the values of $\mathbf{E}_{s\sim\mathcal{U}}[V^{\pi}_{R_E}(s)]$ with $\pi$ being in turn the optimal policy for the rewards given by SCIRL, CSI and RE, the policy $\pi_C$ of the classifier (the very one the classification step of CSI outputs), and the optimal policy for a randomly drawn reward. Performance for CSI is slightly but definitively higher than for SCIRL (see the $p$-values for the mean equality test in Tab.~\ref{tab:highway}), slightly below the performance of the expert itself. Very few samples (100) are needed to reliably achieve expert-level performance.

>>>>

%\usepackage{amsthm}
\usepackage{subfig}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{soul}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage[integrals]{wasysym}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage{algorithmic}
%\usepackage{subfigure}
%\usepackage{natbib}
% scientific notation, 1\e{9} will print as 1x10^9
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}
\usepackage{booktabs} % Fancy tables
\tolerance=1000
\usepackage{amsmath}
%\newtheorem{definition}{Definition}
%\newtheorem{theorem}{Theorem}
%\newtheorem{corollary}{Corollary}
\usepackage{dsfont}
\usepackage{mathtools}
\mathtoolsset{showonlyrefs=true}
\providecommand{\alert}[1]{\textbf{#1}}
\newcommand{\E}{\mathbf{E}}
\newcommand{\argmax}{\operatorname*{argmax}} %\operatorname* pour les op. pouvant admettre des limites...
\newcommand{\Card}{\operatorname*{Card}} %\operatorname* pour les op. pouvant admettre des limites...

@@

We work with finite MDPs \cite{puterman1994markov}, that is tuples $\{S,A,P,R,\gamma\}$. The state space is noted $S$, $A$ is a finite action space, $R\in\mathbb{R}^{S\times A}$ is a reward-function, $\gamma\in (0,1)$ is a discount factor and $P\in \Delta_{S}^{S\times A}$ is the Markovian dynamics of the MDP. Thus, for each $(s,a)\in S\times A$, $P(.|s,a)$ is a distribution over $S$ and $P(s'|s,a)$ is the probability to reach $s'$ by choosing action $a$ in state $s$. At each time step $t$, the agent uses the information encoded in the state $s_t\in S$ in order to choose an action $a_t \in A$ according to a (deterministic\footnote{We restrict ourselves here to deterministic policies, but the loss of generality is minimal as there exists at least one optimal deterministic policy.}) policy $\pi\in A^S$. The agent then steps to a new state $s_{t+1}\in S$ according to the Markovian transition probabilities $P(s_{t+1}|s_t,a_t)$. Given that $P_\pi = (P(s'|s,\pi(s)))_{s,s' \in S}$ is the irreductible transition probability matrix, the stationary distribution over the states $\rho_\pi$ induced by a policy $\pi$ satisfies $\rho_\pi^TP_\pi = \rho_\pi^T\textrm{, with }X^T\textrm{ being the transpose of }X$.
The stationary distribution relative to the exert policy $\pi_E$ is $\rho_E$.

The reward function $R$ is a local measure of the quality of the control. The global quality of the control induced by a policy $\pi$, with respect to a reward $R$, is assessed by the value function $V^\pi_R \in \mathbb{R}^{S}$ which associates to each state the expected discounted cumulative reward for following policy $\pi$ from this state:
$V^\pi_R(s) = \E[\sum_{t\geq 0}\gamma^tR(s_t,\pi(s_t))|s_0 = s,\pi]$.
This long-term criterion is what is being optimized when solving an MDP. Therefore, an optimal policy $\pi^*_R$ is a policy whose value function (the optimal value function $V^*_R$) is greater than that of any other policy, for all states: $\forall \pi, V^*_R\geq V^\pi_R$.

The Bellman evaluation operator $T^\pi_R: \mathbb{R}^{S} \rightarrow  \mathbb{R}^{S}$ is defined by
$  T^{\pi}_RV = R_\pi + \gamma P_\pi V$
where $R_\pi = (R(s,\pi(s)))_{s\in S}$. The Bellman optimality operator follows naturally:
 $ T^*_RV = \max_\pi T^\pi_RV$.
  Both operators are contractions. The fixed point of the Bellman evaluation operator $T^\pi_R$ is the value function of $\pi$ with respect to reward R:
    $V^\pi_R = T^\pi_R V^\pi_R \Leftrightarrow V^\pi_R = R_\pi + \gamma P_\pi V^\pi_R$.
The Bellman optimality operator $T^*_R$ also admits a fixed point, the optimal value function $V_R^*$ with respect to reward $R$.

@@

In this section, we prove that the deterministic expert policy $\pi_E$ is near optimal for the reward $\hat{R}^C$ the regression step outputs.
More formally, recalling from Sec.~\ref{sec:background} that $\rho_E$ is the stationary distribution of the expert policy, we prove that $\E_{s\sim\rho_E}[V^*_{\hat{R}^C}(s)-V^{\pi_E}_{\hat{R}^C}(s)]$ is bounded by a term that depends on:
\begin{itemize}
  \item the classification error defined as
$\epsilon_C=\E_{s\sim\rho_E}[\mathds{1}_{\{\pi_C(s)\neq\pi_E(s)\}}]$;
\item the regression error defined as 
 $   \epsilon_R=\max_{\pi\in A^S}\|\epsilon_\pi^R\|_{1,\rho_E}$,
  with:
  \begin{itemize}
  \item the subscript notation already used for $R_\pi$ and $P_\pi$ in Sec.~\ref{sec:background} meaning that, given an $X\in\mathbb{R}^{S\times A}$, $\pi\in A^S$, and $a\in A$,  $X_\pi\in\mathbb{R}^S$ and $X_a\in\mathbb{R}^S$ are respectively such that: $\forall s\in S, X_\pi(s)=X(s,\pi(s))$ and $\forall s\in S, X_a(s)=X(s,a)$ ;
  \item $\epsilon^R_\pi=R^C_\pi-\hat{R}^C_\pi$ ;
  \item $\|.\|_{1,\mu}$ the $\mu$-weighted $L_1$ norm: $\|f\|_{1,\mu} = \E_{x\sim \mu}[|f(x)|]$;
  \end{itemize}
\item the concentration coefficient $C_* = C_{\hat \pi_C}$ with:
  \begin{itemize}
  \item $C_{\pi}=(1-\gamma)\sum_{t\geq0}\gamma^tc_{\pi}(t), \text{ with } c_{\pi}(t)=\max_{s\in S}\frac{(\rho_E^TP^t_\pi)(s)}{\rho_E(s)}$ ;
  \item $\hat \pi_C$, the optimal policy for the reward $\hat R^C$ output by the algorithm ;
  \end{itemize}
  The constant $C_*$ can be estimated \emph{a posteriori} (after $\hat{R}^C$ is computed). {\it A priori}, $C_*$ can be upper-bounded by a more usual and general concentration coefficient but $C_*$ gives a tighter final result: one can informally see $C_*$ as a measure of the similarity between the distributions induced by $\hat\pi_C$ and $\pi_E$ (roughly, if $\hat{\pi}_C\approx\pi_E$ then $C_* \approx 1$).
\item $\Delta q = \max_{s\in S}(\max_{a\in A}q(s,a)-\min_{a\in A}q(s,a))=\max_{s\in S}(q(s,\pi_C(s))-\min_{a\in A}q(s,a))$,
    which could be normalized to $1$ without loss of generality. The range of variation of $q$ induces the one of $R^C$, $\hat{R}^C$ and $V^{\pi_C}_{\hat{R^C}}$.
  \end{itemize}
\begin{theorem}
\label{thm}
Let $\pi_E$ be the deterministic expert policy, $\rho_E$ its stationary distribution and $\hat{R}^C$ the reward the cascading algorithm outputs. We have:
\begin{equation}
0\leq\E_{s\sim\rho_E}[V^*_{\hat{R}^C}(s)-V^{\pi_E}_{\hat{R}^C}(s)]\leq \frac{1}{1-\gamma}\left(\epsilon_C\Delta q +\epsilon_R(1+C_*)\right).
\end{equation}
\end{theorem}
\begin{proof}
First let's recall some notations, $q\in\mathbb{R}^{S\times A}$ is the score function output by the classification step, $\pi_C$ is a deterministic classifier policy so that $\forall s \in S,\pi_C(s)\in\argmax_{a\in A}q(s,a)$, $R^C\in\mathbb{R}^{S\times A}$ is so that $\forall (s,a)\in S\times A, R^C(s,a)=q(s,a) -\gamma\sum_{s'\in S}P(s'|s,a)q(s',\pi_C(s'))$,
and $\hat{R}^C\in\mathbb{R}^{S\times A}$ is the reward function output by the regression step.

@@

We no have the following vectorial equalities $
R^C_a=q_a-\gamma P_aq_{\pi_C} ;
R^E_a=q_a-\gamma P_aq_{\pi_E};
\epsilon^R_a=R^C_a-\hat{R}^C_a$.
Now, we are going to upper bound the term: $\E_{s\sim\rho_E}[V^*_{\hat{R}^C}-V^{\pi_E}_{\hat{R}^C}]\geq0$ (the lower bound is obvious as $V^*$ is optimal).
Recall that $\hat{\pi}_C$ is a deterministic optimal policy of the reward $\hat{R}^C$. First, the term $V^*_{\hat{R}^C}-V^{\pi_E}_{\hat{R}^C}$ is decomposed:
\begin{equation}
V^*_{\hat{R}^C}-V^{\pi_E}_{\hat{R}^C}=(V^{\hat{\pi}_C}_{\hat{R}^C}-V^{\hat{\pi}_C}_{R^C})+(V^{\hat{\pi}_C}_{R^C}-V^{\pi_E}_{R^C})+(V^{\pi_E}_{R^C}-V^{\pi_E}_{\hat{R}^C}).
\end{equation}
We are going to bound each of these three terms. First, let $\pi$ be a given deterministic policy. We have, using $\epsilon^R = R^C - \hat R^C$:
$V^{\pi}_{R^C}-V^{\pi}_{\hat{R}^C}=V^{\pi}_{\epsilon^R}=(I-\gamma P_\pi)^{-1}\epsilon^R_{\pi}$.
If $\pi=\pi_E$, we have, thanks to the power series expression of $(I-\gamma P_{\pi_E})^{-1}$, the definition of $\rho_E$ and the definition of the $\mu$-weighted $L_1$ norm, one property of which is that $\forall X, \mu^TX \leq \|X\|_{1,\mu}$:
\begin{equation}
  \rho_E^T(V^{\pi}_{R^C}-V^{\pi}_{\hat{R}^C})=\rho_E^T(I-\gamma P_{\pi_E})^{-1}\epsilon^R_{\pi_E}=\frac{1}{1-\gamma}\rho_E^T\epsilon^R_{\pi_E}\leq\frac{1}{1-\gamma}\|\epsilon^R_{\pi_E}\|_{1,\rho_E}.\end{equation}

@@

$\rho_E^T((V^{\pi_E}_{R^C}-V^{\pi_E}_{\hat{R}^C})+(V^{\hat{\pi}_C}_{\hat{R}^C}-V^{\hat{\pi}_C}_{R^C}))\leq\frac{1+C_*}{1-\gamma}\epsilon_R$.

@@

$\rho_E^T(V^{\pi_C}_{R^C}-V^{\pi_E}_{R^E})\leq\Delta q\sum_{s\in S}\rho_E(s)[\mathds{1}_{\{\pi_C(s)\neq\pi_E(s)\}}]= \Delta q \epsilon_C$.

@@

\rho_E^T(V^{\pi_E}_{R^E}-V^{\pi_E}_{R^C})&=\rho_E^T(I-\gamma P_{\pi_E})^{-1}(R^E_{\pi_E}-R^C_{\pi_E})=\rho_E^T(I-\gamma P_{\pi_E})^{-1}\gamma P_{\pi_E}(q_{\pi_C}-q_{\pi_E}),

@@

If we combine all of the results, we obtain the final bound as stated in the theorem.

@@

More precisely, the expert controls the system for $M$ runs of lengths $\{L_i\}_{1\leq i \leq M}$, %with values of $n$ in $\{3,7,10,15,20\}$,
giving samples $\{(s_k,a_k=\pi_E(a_k),s'_k)_k\} = D_E$. The dataset $D_C$ fed to the classifier is straightforwardly constructed from $D_E$ by dropping the $s'_k$ terms:
$  D_C = \{(s_i = s_k,a_i=a_k)_i\}$.
\subsubsection{Heuristics}
\label{subsubsec:heuristics}

There is no particular problem for the classification step. It is however not reasonable to construct the dataset $D_R = \{((s_k,a_k=\pi_E(s_k)),\hat r_k)_k\}$ only from expert transitions and expect a small regression error term $\epsilon_R$. Indeed, the dataset $D_E$ only samples the dynamics induced by the expert's policy and not the whole dynamics of the MDP. 
This means that for a certain state $s_k$ we only know the corresponding expert action $a_k=\pi_E(s_k)$ and the following state $s'_k$ sampled according to the MDP dynamics : $s'_k \sim P(\cdot|s_k,a_k)$. For the regression to be meaningful, we need samples associating the same state $s_k$ and a different action $a\neq a_k$ with a datapoint $\hat r \neq \hat r_k$.\\
Recall that $\hat r_j = q(s_j,a_j) - \gamma q(s'_j,\pi_C(s'_j))$ (Eq.~\eqref{eq:rj}); without knowing $s'_k \sim P(\cdot|s_k,a\neq a_k)$, we cannot provide the regressor with a datapoint to associate with $(s_k,a\neq a_k)$.
We artificially augment the dataset $D_R$ with samples $((s_j=s_k,a),r_{min})_{j;\forall a\neq \pi_E(s_j) = a_k}$ where
$  r_{min} = \min_k\hat r_k - 1$.
This heuristics intructs the regressor to associate a state-action tuple disagreeing with the expert ({\it i.e.}, $(s_k,a\neq a_k)$) with a reward strictly inferior to any of those associated with expert state action tuples ({\it i.e.}, $(s_k,a_k=\pi_E(s_k))$). Semantically, we are asserting that disagreeing with the expert in states the expert visited is a bad idea. This heuristics says nothing about states absent from the expert dataset. For such states the generalization capabilities of the regressor and, later on, the exploration of the MDP by an agent optimizing the reward will solve the problem.
Although this heuristics was not analyzed in Sec.~\ref{sec:analysis} (where the availability of a more complete dataset $D_R$ was assumed), the results shown in the next two subsections demonstrate its soundness.
\subsubsection{Comparison with state-of-the-art approaches}
The similar looking yet fundamentally different algorithm SCIRL \cite{klein2012scirl} is an obvious choice as a contender to CSI as it advertises the same ability to work with very little data, without repeatedly solving the RL problem. In both experiments we give the exact same data to CSI and SCIRL.

@@

The objective criterion for success is the number of steps needed to attain the goal when starting from a state picked at random ; the lesser the better. We can see Fig.~\ref{fig:MountainCar} that the optimal policies for the rewards found by SCIRL and CSI very rapidly attain expert-level performance and outperform the optimal policy for the reward of RE and the classification policy. When very few samples are available, CSI does better than SCIRL (with such a low $p$-value for $n=10$, see Tab.~\ref{tab:mc}, the hypothesis that the mean performance is equal can be rejected); SCIRL catches up when more samples are available. Furthermore, CSI required very little engineering as we cascaded two off-the-shelf implementations whereas SCIRL used hand-tuned features and a custom classifier.
\begin{table}
  \small
  \subfloat[Mountain Car\label{tab:mc}]{
    \begin{tabular}{cc}
      \toprule
      Number of expert samples & $p$-value\\
      \midrule
      $ 10$ & $\mathbf{1.5e-12}$\\
      $ 30$ & $3.8e-01$\\
      $100$ & $1.3e-02$\\
      $300$ & $7.4e-01$\\
      \bottomrule
    \end{tabular}
  }
  \hfill
  \subfloat[Highway Driving\label{tab:highway}]{
    \begin{tabular}{cc}
      \toprule
      Number of expert samples & $p$-value\\
      \midrule
      $  9$ & $3.0e-01$\\
      $ 49$ & $\mathbf{8.9e-03}$\\
      $100$ & $\mathbf{1.8e-03}$\\
      $225$ & $\mathbf{2.4e-05}$\\
      $400$ & $\mathbf{2.0e-50}$\\
      \bottomrule
    \end{tabular}
  }
  \caption{Student or Welch test of mean equality (depending on wether a Bartlett test of variance equality succeeds) $p$-values for CSI and SCIRL on the mountain car (\ref{tab:mc}) and the highway driving simulator (\ref{tab:highway}). High values ($>1.0\e{-02}$) means that the hypothesis that the means are equal cannot be rejected.}
\end{table}
\begin{figure}
  \center
  \includegraphics[width=.71\textwidth]{Exp11.pdf}
  \caption{Performance of various policies on the mountain car problem. This is the mean over 100 runs.}
  \label{fig:MountainCar}
\end{figure}

  \begin{figure}
   \subfloat[Mean performance over 100 runs on the Highway driving problem.]{\label{tutu}\includegraphics[width=.5\linewidth]{"Exp14"}}
   \subfloat[Zoom of Fig~\ref{tutu} showing the ranking of the three IRL algorithms.]{\includegraphics[width=.5\linewidth]{"Exp14_zoom"}}
  \caption{Results on the highway driving problem.}
  \label{fig:Highway}

@@

Results are shown Fig.~\ref{fig:Highway}. We give the values of $\mathbf{E}_{s\sim\mathcal{U}}[V^{\pi}_{R_E}(s)]$ with $\pi$ being in turn the optimal policy for the rewards given by SCIRL, CSI and RE, the policy $\pi_C$ of the classifier (the very one the classification step of CSI outputs), and the optimal policy for a randomly drawn reward. Performance for CSI is slightly but definitively higher than for SCIRL (see the $p$-values for the mean equality test in Tab.~\ref{tab:highway}, from 49 samples on), slightly below the performance of the expert itself. Very few samples (100) are needed to reliably achieve expert-level performance.
